---
title: "Intro to Quarto" #yaml section
format: html
bibliography: references.bib
---

# 1. Introduction

<!-- -->

## a. About the data

-   water sample data from the Arctic ocean
-   data downloaded from the Arctic Data Center: [Tweedie ADC data](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## b. Setup

```{r}
#| warning: false

library(readr)
library(here)  # helpful for relative paths

```

## c. Read in data

```{r}
#| warning: false

bg_chem <- read_csv(here("data/BGchem2008data.csv"))

```

# 2. Analysis

<!-- -->

```{r}
#| eval: false
#| echo: false  # include source code in the output


# print column names from dataset
colnames(bg_chem)

# show data structure
str(bg_chem)

#show first 6 lines of the dataframe
head(bg_chem)




# print unique value
unique(bg_chem$Si)


```

## a. Calculate summary statistics

```{r}
# get summary of the dataset
summary(bg_chem)
```


## b. Calculate mean Redfield ratio

```{r}
#calculate some means of nitrate and nitrite
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)

amn <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```


## c. Plot Redfield ratio

```{r}
ratio <- (nitrate + nitrite + amn) / phos
```

Here is the ratio we calculated: `r round(ratio)` 

# 3. Conclusion

For future reference

Markdown formatting

# This is a large title

## this is a smaller title

### [@chawarski2022]

| Col1 | Col2 | Col3 |
|------|------|------|
| x    | y    | z    |
| 2    | 4    | 2    |
| 5    | 7    | 9    |

### this is the smallest title

*text*

-   list of cool stuff
    -   cats
    -   dogs
    -   skateboards
    -   
