---
title: "Tidying and Wrangling"
format: html
---

```{r, echo=FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

# Data Sources

## Read in our data

```{r, message = FALSE}

catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

```

## Explore

```{r}
colnames(catch_original)

tail(catch_original)


```

## Repair data

```{r}
catch_modified <- catch_original
catch_modified$Chinook <- as.numeric(catch_modified$Chinook)


```

# Cleand & Reshaping

## using the select() function
```{r}
# select() to keep columns that we want
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
head(catch_data)

# to 


```

## changing column contents with mutate()

```{r}
catch_clean <- catch_original %>% 
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = as.numeric(Chinook))
```

```{r}
is.na(catch_clean$Chinook)

# get the location of the NA value

i <- which(is.na(catch_clean$Chinook))

# look at that row in the data
catch_clean[i, ]

# use mutate to fix
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(condition = Chinook == "I",
                           true = "1", #if true, change to 1 
                           false = Chinook), # if false, take original value from Chinook column
         Chinook = as.numeric(Chinook))
    
head(catch_clean)

catch_clean[i,]

```


## changing file shape using pivot_longer() and pivot_wider()

```{r}

catch_long <- catch_clean %>% 
  pivot_longer(cols = c(Chinook, Sockeye, Coho, Pink, Chum),
               names_to = "species", 
               values_to = "catch")

catch_wide <- catch_long %>% 
  pivot_wider(names_from = species,
              values_from = catch)

```

## rename columns with rename()

```{r}
catch_long <- catch_long %>% 
  rename(catch_thousands = catch)
head(catch_long)
```


## use mutate to create a new column

```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)

head(catch_long)
```

## summary statistics using group_by and summarise()

```{r}
# get average catch by region
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))

n_region <- catch_clean %>% 
  group_by(Region) %>% 
  summarise(n = n())

```

# Filter rows using filter()

```{r}
sse_catch <- catch_long %>% 
  filter(Region == "SSE")


catch_million <- catch_long %>% 
  filter(catch > 1000000)

```

## Sorting data with arrange()
```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))
```

## Combining dataset together with join()


```{r}
region_defs <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)

catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)


```


## split and combine values in columns

```{r}
sites_df <- data.frame(site = c("HAW-101", "HAW-102", "OAH-320"))
head(sites_df)

split_df <- sites_df %>% 
  separate(site, c("island", "site_numer"), sep = "-")
```
```{r}
dates_df <- data.frame(
  year = c("1930", "1930", "1932"),
  month = c("12","12", "12"), 
  day = c("14", "15", "16")
)

dates_unite <- dates_df %>% 
  unite(col = date, year, month, day, sep = "-")

```



```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), 
               names_to = "species", 
               values_to = "catch") %>%
  mutate(catch = catch * 1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))
```

## save our results with write_csv()
```{r}
library(here)
write_csv(mean_region, here("data/mean_catch_by_region.csv"))


```



